---
title: "Sampling GAML model and storing in data.frame"
output: html_notebook
---
## Setting up GAMA 

R notebook to use R in relation with sampling GAMA. Functions are: 1) automatic creating XML Experiment files, 2) communication with GAMA, 3) Automatic import of experiment results in R data.frame 4) Export to CSV file. 

First, libraries of gamar and XML are imported, and the Installation of GAMA is set up. Note that the setup GAMA gives an error for a version of GAMA later than 1.8, but this does not harm functionality. 

```{r}
# Makes use of the following libraries
library(gamar)
library(XML)

# link gamar to preferred GAMA installation
setup_gama("/home/daan/GAMA");
```

## Loading XML creation function
Secondly, we have to load the function that automatically creates XML files that feed into the call_gama() function from the gamar library. The XML file provides information for the Java command to launch a GAMA experiment. To make an experiment XMl file, we need the following information: 

---
* `single_params`; A single-line data.frame with the input parameters for one model run. 
* `obs`; single-line data.frame with observables
* `simulation_id`; the number of the simulation, relevant for running multiple replications
* `model_path`; the path to the .gaml file containing the model
* `exp_name`; the name of the experiment in the .gaml file
* `until`; till which cycle the model should run

The function puts out the temporary path name to the generated XML file. 


```{r}
# Function creating temporary XML to run the experiment 
createXML <- function(single_params,
                      obs,
                      simulation_id,
                      model_path,
                      exp_name,
                      until) {
  tmp_xml = tempfile(fileext = ".xml")
  
  # Empty XML tree
  xml<-xmlTree() # Empty XML tree
  xml$addTag("Experiment_plan",close=FALSE) # Experiment layer
  xml$addTag("Simulation",close=FALSE, attrs=c(id=paste0(simulation_id), sourcePath=paste0(model_path), finalStep=paste0(until),experiment=paste0(exp_name))) # Simulation layer
  xml$addTag("Parameters",close=FALSE) # Parameters layer
  
  # Add all paramaters and values 
  for (i in 1:ncol(single_params)) {
    xml$addTag("Parameter",attrs=c(name=colnames(single_params)[i],type="FLOAT",value=single_params[1,i]))
  }
  xml$closeTag()
  
  xml$addTag("Outcomes",close=FALSE) # Outcome layer
  
  # Add all outcomes
  for (i in 1:ncol(obs)) {
    xml$addTag("Output",attrs=c(id = paste0(i), name=colnames(obs)[i],framerate=obs[1,i]))
  }
  
  xml$closeTag()
  xml$closeTag()
  xml$closeTag()
  
  saveXML(xml,file=tmp_xml)
  
  return(tmp_xml)
}
```

## Initiating Parametres
Avoiding setup of experiments using gamar, we have to manually define parameters spaces and stepsize. Note that these must correspond with the parameters as defined in the experiment in the GAMA model. 

```{r}
# define to be researched parametres as c(min,max,stepsize) CHECK IF (max-min)%stepsize == 0 !

params = data.frame("alpha"=c(0.1,0.1,0.4),"beta"=c(0,1.0,1.0), "ParallelServed"=c(1,5,4), "gamma"=c(1,11,5),"epsilon"=c(0,1.0,0.5))  # Queue testing
# params = data.frame("ParallelServed"=c(1,10,1), "gamma"=c(0,30.0,10)) # Facility testing

# params = data.frame("alpha"=c(0.0,1.0,1.0),"beta"=c(0,1.0,1.0),"avg_pc"=c(0,1.0,1.0)) # network testing for extreme values
# params = data.frame("alpha"=c(0.1,0.5,0.4),"beta"=c(0,1.0,0.5)) # network testing sampling

#params = data.frame("alpha"=c(0.1,0.1,1.0),"gamma"=c(0.0,30.0,10.0)) # individual household testing


# create full list of all the to be researched entries 
params_list <- vector(mode="list", length=ncol(params))
# add names to list
names(params_list) <- names(params)

for (p in (1:ncol(params))){ # create list
  min <- params[1,p]
  max <- params[2,p]
  step <- params[3,p]
  
  tmplist <- c()
  for (i in 0:((max-min)/step)) {
    tmplist <- append(tmplist,min+step*i)
  }
  params_list[[p]] <- tmplist
} 

```

Next, we have to make a permutation of the input parameter list as the basis for an exhaustive experiment (all combinations of parameters). 

```{r}
# Create samples for all variables  !! not automatic 
params_sample <- expand.grid(params_list$alpha,params_list$beta,params_list$ParallelServed,params_list$gamma,params_list$epsilon) # Queue testing
# params_sample <- expand.grid(params_list$ParallelServed,params_list$alpha,params_list$beta,params_list$gamma)

# params_sample <- expand.grid(params_list$alpha,params_list$beta,params_list$avg_pc) # network testing extreme values
# params_sample <- expand.grid(params_list$alpha,params_list$beta) # network testing sampling

# params_sample <- expand.grid(params_list$alpha,params_list$gamma) # individual household testing 
names(params_sample) <- names(params)
```

Next, we have to set our observables in a data.frame. The names must correspond with the observations as defined in the experiment in the GAMA model. The value it equals the framerate: framerate of 10 means fetching the output values every 10 ticks. 

```{r}
# create dataframe for observations and corresponding framerate 
# obs = data.frame("tick"=1,"AverageEmotionalState"=1,"EmotionalState"=1) # Network testing 
# obs = data.frame("tick"=1,"EmotionalState"=1,"FoodInStorage"=1) # Individual household testing
# obs = data.frame("tick"=1,"AverageDeliverySize"=1,"QueueLength"=1,"FacilityStorage"=1)
obs = data.frame("tick"=1,"AverageDeliverySize"=1,"QueueLength"=1,"FacilityStorage"=1,"AverageEmotionalState"=1)

```

Next, we set the input parameters that will feed into the `createXML` function

```{r}
# set path to model file 
path = "/home/daan/GAMA/workspace/FATM_ABM/main/facility_verification/QReaction_test.gaml"
#path = "/home/daan/GAMA/workspace/FATM_ABM/main/facility_verification/facilities_test.gaml"

# path = "/home/daan/GAMA/workspace/FATM_ABM/main/household_verification/household_test.gaml"
# path = "/home/daan/GAMA/workspace/FATM_ABM/main/household_verification/network_test.gaml"

# give experiment name 
# exp_name = "NoStartle"
# exp_name = "NetworkStartle"
exp_name = "FacilityToHouseholds"

# till what tick 
cycles_in_day <- 6*24
days_runtime<-30
until = paste0(cycles_in_day*days_runtime)

replications <- 1

```


## Running the model

Lastly, we initialise the data.frame in which all results + input parametres will be stored as NULL. For every row in the permutation list `params_sample` and for the number of replications `replications` the model is initialised and the results defined in `obs` are stored in the `total_df`. 

```{r}
tb <- Sys.time()

# empty df 
total_df <- NULL

# for the entire sample list 
for (run in 1:nrow(params_sample)){
  for (replication in 1:replications){
    simulation_id<-paste0(replication)
    
    # temporary experiment xml
    tmp_xml <- createXML(single_params = params_sample[run,],obs=obs,simulation_id = simulation_id,model_path = path,exp_name =   exp_name,until=until)
    tmp_dir <- tempdir() # temporary dir to store outcome xml
    message("call gama model")
  
    call_gama(tmp_xml,hpc=2,output_dir = tmp_dir) # run gama model
    
    message("import outcome xml")
    # import outcome xml 
    tmp_df<-XML::xmlToDataFrame(XML::xmlParse(paste0(tmp_dir,"/","simulation-outputs",simulation_id,".xml")), stringsAsFactors = FALSE)
    # add column names
    names(tmp_df)<-colnames(obs)
    
    # delete temporary xml files
    unlink(paste0(tmp_dir,"/","simulation-outputs",simulation_id,".xml"))
    unlink(tmp_xml)
    
    # add input parameters to df
    for (p in 1:ncol(params_sample)){
      tmpcol <- rep(params_sample[run,p],until)
      
      tmp_df <- cbind(tmpcol,tmp_df)
      names(tmp_df)[1] <- names(params_sample)[p]
    }
    
    tmpcol <- rep(simulation_id,until)
    
    tmp_df <- cbind(tmpcol,tmp_df)
    names(tmp_df)[1] <- "Replication"
  
    # add to existing dataframe 
    if (is.null(total_df)){
      total_df<-tmp_df
    } else{
      total_df <- rbind(total_df, tmp_df)
    }
  }
}


# Write results to csv
write.csv(total_df,paste0("/home/daan/Desktop/",exp_name,".csv"))

ta <- Sys.time()

message("took me ",ta-tb)
```

Now we can take a look at the results:

```{r}
print(total_df)
```


